---
title: "AC584 Report"
subtitle: "Second Semester R Studio Assignment"
author: "Cianan Markey"
date: "2024-04-04"
html: 
toc: true 
theme: sketchy
backgroundcolor: "lightgrey"

execute: 
  echo: false
  warning: false
  message: false
editor_options: 
  chunk_output_type: console
---
```{r}
#| label: setup
#| include: false

# libraries
library(tidyverse)
library(plotly)
library(scales)

# data
unicef_indicator_1 <- read_csv("unicef_indicator_1.csv")
unicef_metadata <- read_csv("unicef_metadata.csv")
unicef_indicator_2 <- read_csv("unicef_indicator_2.csv")
continents <- read_csv("continents.csv")

```
## Introduction ##

Welcome to my report. I have created this report to investigate the data provided to me from **Unicef** regarding Anaemia infection rates and more general statistics like *GDP, Life Expectancy and Population*.

## Anaemia Infection Rates Map ##
This Map shows us the anaemia infection rates of the given countries.

<span style="color: red;"> I utilised the scale fill gradient code in order to enable me to create a gradient which highlights the change in the infection rates in the different countries

[This **link** is where I have received my datasets from](https://data.unicef.org/dv_index/)


[There is **more information** about my indicator (anaemia) and many other ***indicators*** available here on this website](https://data.unicef.org/indicator-profile/)

```{r}

# MAP #

map_world <- map_data("world")

color_scale <- scale_fill_gradientn(
  colors = c("darkred", "red", "coral"),  
  values = rescale(c(0, 40, 60, 85)),  
  limits = c(0, 85),  
  name = "Indicator"
)

map_data_join <- full_join(unicef_indicator_1, map_world, by = c("country" = "region"))

map_Object <- ggplot(map_data_join, aes(long, lat, group = group, fill = obs_value)) +
  geom_polygon(color = "black") +
  color_scale + 
  labs(x = "", y = "", title = "Anaemia Infection Rates") +
  theme(text = element_text(family = "Times New Roman"))

ggplotly(map_Object)
```
^1: If you hover over one country it will tell you the infection rate for that country.


## Evolution of Life Expectancy Per Continent ##
This time series visualisation enables us to see the progression in life expectancy in the different continents over decades.

Each **continent** is linked to a *colour* as follows:
 <span style="color: blue;"> Asia,<span style="color: yellow;"> Europe, <span style="color: red;"> Africa,<span style="color: orange;"> Americas,<span style="color: green;"> Oceania
```{r}
# Time Series #

data_join_Continents <- full_join(continents, unicef_metadata, by = join_by(country))

timeseries_plot_1 <- data_join_Continents %>%
  na.omit() %>%
  ggplot() +
  aes(time_period, lifeExp, group = country, color = continent) +
  geom_line() +
  scale_color_manual(values = c("Asia" = "blue", 
                                "Europe" = "yellow",
                                "Africa" = "red",
                                "Americas" ="orange",
                                "Oceania" = "green")) + 
  labs(x = "Life Expectancy", y = "Year", title = "Evolution of Life Expectancy") +
  theme_classic() + 
  theme(text = element_text(family = "Times New Roman"))


ggplotly(timeseries_plot_1)
```
<span style="color: #FFFF00;"> ^1: If you click on the continent in the legend it will remove that continent from the visualisation

<span style="color: #FF00FF;"> ^2: If you double click on one continent in the legend it will highlight that continent.


## Continental Scatter Plots ##
***This visualisation enables us to see the relationship between GDP and Life Expectancy in each continent***

<span style="text-decoration: underline;">Benefits of this Visualisation</span>

- Continental Comparison
- Colour Co-Ordination with Time Series
- Use of Facet Wrap

```{r}
# Scatter Plot #

data_scatterplot_1 <- data_join_Continents %>%
  na.omit()%>%
  ggplot() +
  aes(gdp, lifeExp, color = continent, size = population) +
  geom_point(alpha = 0.5) +
  facet_wrap(~ continent) +
  scale_color_manual(values = c("Asia" = "blue", 
                                "Europe" = "yellow",
                                "Africa" = "red",
                                "Americas" ="orange",
                                "Oceania" = "green")) +
  scale_x_continuous(limits = c(0, 75000), breaks = c(25000, 50000, 75000), labels = scales::unit_format(unit = "K", scale = 0.001)) + 
  labs(x = "GDP per Capita in USD", y = "Life Expectancy at Birth", title = "Continental Correlation Between Life Expectancy and GDP") +
  theme_light() + 
  theme(text = element_text(family = "Times New Roman"))

ggplotly(data_scatterplot_1)
```

## Population Bar Charts 1961 v 2021 ##

*These bar charts enable us to see the growth in population of the five continents over the 60 year period from 1961-2021*

This first graph shows us the **total population** per continent in <span style="color: green;"> ***1961***

```{r}

# Bar Chart 1961 #

data_1961 <- subset(data_join_Continents, time_period == 1961)

total_population_1961 <- aggregate(population ~ continent, data = data_1961, FUN = sum)

BarChart_1961 <- ggplot(total_population_1961, aes(x = reorder(continent,population), y = population, fill = continent)) +
  geom_bar(stat = "identity") +
  labs(title = "Total Population by Continent in 1961",
       x = "Continent", y = "Total Population")+
  scale_y_continuous(labels = scales::comma)+
  theme_classic() + 
  theme(text = element_text(family = "Times New Roman"))

ggplotly(BarChart_1961)

```

```{r}

# Bar Chart 2021 #

data_2021 <- subset(data_join_Continents, time_period == 2021)

total_population <- aggregate(population ~ continent, data = data_2021, FUN = sum)

BarChart_2021 <- ggplot(total_population, aes(x = reorder(continent,population), y = population, fill = continent)) +
  geom_bar(stat = "identity") +
  labs(title = "Total Population by Continent in 2021",
       x = "Continent", y = "Total Population")+
  scale_y_continuous(labels = scales::comma)+
  theme_classic() + 
  theme(text = element_text(family = "Times New Roman"))

ggplotly(BarChart_2021)

```
**Notice** how the Y-axis has ***drastically*** increased in value over the 60 years. 

This second graph shows us the **total population** per continent in <span style="color: blue;"> ***2021***

Similarly to the first graph, the **continent** with the highest population is <span style="color: Green;"> Asia.

However the **continent** which has seen the biggest increase in their population is <span style="color: Red;"> Africa.


## Conclusion ##

<span style="color: #FFA500;"> Thank you for taking the time to engage with my report. <span style="color: #00FF00;"> I hope you found it <span style="color: #00FFFF;"> engaging and interesting. 

